---
title: "FLS Unit 3"
author: "Justin Ehly"
date: "September 5, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
fifa <- read.csv("C:/Users/justi/Google Drive/_SMU/6306/MSDS_6306_Doing-Data-Science/Unit 3/FIFA Players.csv")

unit3q1 = fifa %>%
  filter(Position == "LM" | Position == "LF")
unit3q1$Position = as.character(unit3q1$Position)
str(unit3q1$Position)
unit3q1

unique(unit3q1$Position)

unit3q1 %>%
  select(Position, Acceleration, Agility) %>%
  ggpairs(aes(color = Position))

```

# R code for drawing and shading

```{r}
#' Credit: Volodymyr Orlov
#' modified by MSDS SMU
#' https://github.com/VolodymyrOrlov/MSDS6371/blob/master/shade.r
#' Draws a t-distribution curve and shades rejection regions
#' 
#' @param df degrees of freedom.
#' @param alpha significance level
#' @param h0 null hypothesis value
#' @param sides one of: both, left, right
#' @param t_calc calculated test statistics
#' @examples
#' shade(49, 0.05, 0, t_calc=1.1)
#' shade(91, 0.05, 0, t_calc=NULL, sides = 'right')
#' shade(7, 0.05, 0, t_calc=1.5, sides = 'left')
#' shade(7, 0.05, 0, t_calc=1.5, sides = 'both')

shade <- function(df, alpha, h0 = 0, sides='both', t_calc=NULL) {
  e_alpha = alpha
  if(sides == 'both'){
    e_alpha = alpha / 2
  }
  cv = abs(qt(e_alpha, df))
  curve(dt(x, df), from = -4, to = 4, ylab='P(x)', xaxt='n') 
  abline(v = 0, col = "black", lwd = 0.5)
  labels = h0
  at = 0
  if(sides == 'both' | sides == 'left'){
    x <- seq(-4, -abs(cv), len = 100) 
    y <- dt(x, df)
    polygon(c(x, -abs(cv)), c(y, min(y)), col = "blue", border = NA)
    lines(c(-cv, -cv), c(0, dt(-cv, df)), col = "black", lwd = 1)
    text(-cv - (4 - cv) / 2, 0.05, e_alpha)
    labels = c(round(-cv, 3), labels)
    at = c(-cv, at)
  }
  if(sides == 'both' | sides == 'right'){
    x <- seq(abs(cv), 4, len = 100)
    y <- dt(x, df)
    polygon(c(abs(cv), x), c(min(y), y), col = "blue", border = NA)
    lines(c(cv, cv), c(0, dt(cv, df)), col = "black", lwd = 1)
    text(cv + (4 - cv) / 2, 0.05, e_alpha)
    labels = c(labels, round(cv, 3))
    at = c(at, cv)
  }
  if(is.numeric(t_calc)){
    abline(v = t_calc, col = "red", lwd = 2)
    text(t_calc + 0.5, 0.2, t_calc, col = "red")
  }
  axis(1, at=at, labels=labels)
}
#The above defines the function shade. To use it, you must call it. More examples are in the comments above.

# shade <- function(df, alpha, h0 = 0, sides='both', t_calc=NULL)
shade(14.48, 0.05, 0, t_calc=2.0126)

```

#T-test U3Q4 2-sample
```{r}
LF = fifa %>% filter(Position =="LF")
LM = fifa %>% filter(Position =="LM")
count(LF)

t.test(x=LF$Agility, y=LM$Agility, mu = 0, conf.int = "TRUE", alternative = "two.sided")

```

# Histogram, Boxplots, Points of fifa dataset
```{r}

LM %>%
  ggplot(aes(x = Agility)) +
  geom_histogram(fill = "blue") +
  ggtitle("Left Middlefielders") +
  theme_bw()

LF %>%
  ggplot(aes(x = Agility)) +
  geom_histogram(fill = "blue") +
  ggtitle("Left Forwards") +
  theme_bw()

unit3q1 %>%
  ggplot(aes(y = Agility)) +
  geom_boxplot(aes(fill = Position)) +
  ggtitle("Left Middlefielders vs Left Forwards") +
  theme_bw()


```

# f-test for equal variance
```{r}

var.test(x = LM$Agility, y = LF$Agility, alternative = NULL , ratio = 1, conf.level = 0.95)

```

# cut function
# displ_Factor = cut(mpg$displ, breaks = c(1,4,6,8), labels = c("Low", "Medium", "High"))
# --cut cuts numerical variables are certain thresholds
# --breaks here wer have 4 breaks and that is going to give us levels

# Question 2 Select/create at least 2 categorical variables and select two continuous variables and perform an EDA.  Also, at least one of the categorical variables should be created from a continuous variable (using the cut() function).   

```{r}
fifa <- read.csv("C:/Users/justi/Google Drive/_SMU/6306/MSDS_6306_Doing-Data-Science/Unit 3/FIFA Players.csv")


unit3q2 <- fifa
unit3q2$Wage <- as.numeric((unit3q2$Wage))

 
fifa %>%
  ggplot(aes(Position, fill = Position)) +
  geom_histogram(stat = "count") + 
  ggtitle("Number of Players in Each Position") +
  theme_bw()

WageBreakdown <- fifa %>% filter(Position == "CM" | Position == "CB")
WageBreakdown$Position <- as.character(WageBreakdown$Position)
WageBreakdown$Wage <- as.numeric(WageBreakdown$Wage) 

WageBreakdown %>%
  ggplot(aes(y = Wage)) +
  geom_boxplot() + 
  ggtitle("Wage Breakdown") +
  theme_bw() +
  scale_y_continuous(breaks = seq(0, max(WageBreakdown$Wage), by = 10))

summary(unit3q2$Wage)


unit3q2$WageFactor <- cut(unit3q2$Wage, breaks = c(0,(32+56.42)/2,(56.42+84)/2,144), 
                          labels = c("Low", "Medium", "High"))
  
unit3q2 <- unit3q2 %>% filter(Position == "CM" | Position == "CB")     

unit3q2$Position <- as.character(unit3q2$Position)  

unit3q2 %>%
  select(Position, WageFactor, Aggression, Penalties) %>%
  ggpairs(aes(color = Position)) +
  ggtitle("Position, WageFactor, Aggression, Penalties")

unit3q2$Aggression <- as.numeric(unit3q2$Aggression) 

# Q1: CB's more aggressive than CM's
unit3q2 %>%
  group_by(Position) %>%
  dplyr::summarize(median = median(Aggression))

unit3q2 %>%
  ggplot(aes(x = WageFactor, y = Aggression, color = Position)) +
  geom_point() +
  theme_bw() +
  ggtitle("Aggression by WageFactor by Position")


# Q2: what is the relationship between aggression and penalties?
unit3q2 %>%
  group_by(Position) %>%
  dplyr::summarise(median = median(Penalties))

# Q3: Does wage have anything to do with aggress?
unit3q2 %>%
  group_by(WageFactor) %>%
  dplyr::summarise(median = median(Aggression))

unit3q2 %>%
  group_by(WageFactor, Position) %>%
  ggplot(aes(x = Penalties, y = Aggression, color = Position)) +
  geom_point() +
  theme_bw() +
  ggtitle("Aggression v Penalties")

CB <- unit3q2 %>% filter(Position == "CB")
CM <- unit3q2 %>% filter(Position == "CM")


var.test(x = CB$Wage, y = CM$Wage, alternative = NULL, ratio = 1, conf.level = .95)
t.test(x = CB$Wage, y = CM$Wage, mu = 0, alternative = "greater")

unit3q2 %>%
  group_by(Position) %>%
  filter(Position == "CB" | Position == "CM") %>%
  ggplot(aes(x = Penalties, color = Position)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Penalties by Position")


